character histograms of a HCAndersen text, character transition, and
random text generation using a 1st order Markov model. It should be
tested using a simple test set and with a complicated test set by
recalculating the histograms and transitions for the generated text
and comparing with the original.



\begin{enumerate}
\item Write a program, that reads an UTF-8 text from a file as a
  string, and which has the following type:
  \begin{quote}
    \mbox{\lstinline!readText : filename:string -> string!}
  \end{quote}
\item Write a program that converts a string, such that all letters
  are converted to lower case, and which removes all characters except
  a\ldots z. Is should have the following type:
  \begin{quote}
    \mbox{\lstinline!convertText : src:string -> string!}
  \end{quote}
\item Write a program that counts occurences of each lower-case letter
  of the English alphabet in a string and returns a list. The first
  element of the list should be the count of 'a's, second the count of
  'b's etc. The function must have the type:
  \begin{quote}
    \mbox{\lstinline!histogram : src:string -> character list!}
  \end{quote}
\item Write a program that counts occurences of each pairs of lower-case letter
  of the English alphabet in a string and returns a list of lists. The first
  list should be the count of 'a' followed by 'a's, 'b's, etc., second
  list should be the count of 'b' followed by 'a's, 'b's, etc. etc.
  \begin{quote}
    \mbox{\lstinline!cooccurence : src:string -> character list list!}
  \end{quote}
\end{enumerate}
Der skal også laves:
\begin{itemize}
\item En rapport (maks 2 sider)
\item Unit-tests
\item Implementation skal kommenteres jævnfør kommentarstandarden for F\#
\end{itemize}
